Modeling costs and benefits of genomic wastewater surveillance

### 2023-05-05 paper/model-documentation.qmd
adding model diagram

---
title: Modeling costs and benefits of genomic wastewater surveillance
subtitle: Supplementary appendix draft
#author:
#  - name: Pedro Nascimento de Lima
#    email: plima@rand.org
#    affiliations: 
#        - id: rand
#          name: RAND Corporation
#          department: Engineering and Applied Sciences
#          address: 1776 Main Street
#          city: Santa Monica
#          state: CA
#          postal-code: 90401
#    attributes:
#        corresponding: true
#  - name: and colleagues
#    email: others@rand.org
#    affiliations:
#        - ref: rand
abstract: |
  Introduces a model to assess the value provided by wastewater genomic surveillance network. This document is meant to be used as a supplementary appendix of the paper.
keywords: 
  - Genomic Surveillance
  - Pandemic Response
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      #name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
  elsevier-html:
    keep-tex: true
    journal:
      #name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
  docx:
      cite-style: super
---

@@

where $\dot{\mathbf{S}}, \dot{\mathbf{E}}, \dot{\mathbf{P}}, \dot{\mathbf{I}}, \dot{\mathbf{R}}$ represent the time derivatives of each disease status, greek letters represent progression rates and $p_x\mathbf{m}_x$ terms represent migration flows $\mathbf{m}_x$ modulated by propensity to travel $p_x$ at disease state $x \in \{S, E, P, I, R\}$. Force of infection $\boldsymbol{\lambda}_t$ is computed using a time-varying effective contact-rate vector $\boldsymbol{\beta}_t$ which will depend on contact rates within each strata as well as nonpharmaceutical interventions (discussed later). Infectiousness of each disease status $\mathbf{P}$ and $\mathbf{I}$ are modulated by constants $c_p$ and $c_i$.

>>>>

---
title: Modeling costs and benefits of genomic wastewater surveillance
subtitle: Supplementary appendix draft
#author:
#  - name: Pedro Nascimento de Lima
#    email: plima@rand.org
#    affiliations: 
#        - id: rand
#          name: RAND Corporation
#          department: Engineering and Applied Sciences
#          address: 1776 Main Street
#          city: Santa Monica
#          state: CA
#          postal-code: 90401
#    attributes:
#        corresponding: true
#  - name: and colleagues
#    email: others@rand.org
#    affiliations:
#        - ref: rand
abstract: |
  Introduces a model to assess the value provided by wastewater genomic surveillance network. This document is meant to be used as a supplementary appendix of the paper.
keywords: 
  - Genomic Surveillance
  - Pandemic Response
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      #name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
  elsevier-html:
    keep-tex: true
    mermaid:
      theme: neutral
    journal:
      #name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
  docx:
      cite-style: super
  html:
    keep-tex: true
    mermaid:
      theme: neutral
---

@@

where $\dot{\mathbf{S}}, \dot{\mathbf{E}}, \dot{\mathbf{P}}, \dot{\mathbf{I}}, \dot{\mathbf{R}}$ represent the time derivatives of each disease status, greek letters represent progression rates and $p_x\mathbf{m}_x$ terms represent migration flows $\mathbf{m}_x$ modulated by propensity to travel $p_x$ at disease state $x \in \{S, E, P, I, R\}$. @fig-model-diagram illustrates the model.

```{mermaid}
%%| label: fig-model-diagram
%%| fig-cap: Model diagram
%%| fig-width: 6.5
%%| fig-align: center
flowchart LR
  S --> E --> P --> I --> R
  I --> l --> E
```

Force of infection $\boldsymbol{\lambda}_t$ is computed using a time-varying effective contact-rate vector $\boldsymbol{\beta}_t$ which will depend on contact rates within each strata as well as nonpharmaceutical interventions (discussed later). Infectiousness of each disease status $\mathbf{P}$ and $\mathbf{I}$ are modulated by constants $c_p$ and $c_i$.
